---
title: "Assignment 7: GLMs (Linear Regressios, ANOVA, & t-tests)"
author: "Christina Li"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on generalized linear models. 

## Directions
1. Change "Student Name" on line 3 (above) with your name.
2. Work through the steps, **creating code and output** that fulfill each instruction.
3. Be sure to **answer the questions** in this assignment document.
4. When you have completed the assignment, **Knit** the text and code into a single PDF file.
5. After Knitting, submit the completed exercise (PDF file) to the dropbox in Sakai. Add your last name into the file name (e.g., "Fay_A06_GLMs.Rmd") prior to submission.

The completed exercise is due on Tuesday, March 2 at 1:00 pm.

## Set up your session 
1. Set up your session. Check your working directory. Load the tidyverse, agricolae and other needed packages. Import the *raw* NTL-LTER raw data file for chemistry/physics (`NTL-LTER_Lake_ChemistryPhysics_Raw.csv`). Set date columns to date objects.

```{r, echo = F}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=65),tidy=TRUE)
```

2. Build a ggplot theme and set it as your default theme.

```{r setup}
# getwd()
library(tidyverse)
library(agricolae)
library(ggplot2)
library(lubridate)
library(htmltools)

ChemPhy <- read.csv("./Data/Raw/NTL-LTER_Lake_ChemistryPhysics_Raw.csv", stringsAsFactors = TRUE)

ChemPhy$sampledate <- as.Date(ChemPhy$sampledate, 
                              format = "%m/%d/%y")

#2
Theme_1 <- theme_bw (base_size = 11)+
  theme (
    plot.title = element_text(face = "bold", 
                              size = 12, 
                              hjust=0.5),
    axis.text.x = element_text(vjust = 1, 
                               hjust = 0, 
                               size = 10),
    legend.margin = margin(1),
    legend.position = "right",
    legend.text = element_text(size = 10), 
    legend.title = element_text(size = 10))

theme_set(Theme_1)

```

## Simple regression
Our first research question is: Does mean lake temperature recorded during July change with depth across all lakes?

3. State the null and alternative hypotheses for this question:
> Answer:
H0: The mean lake temperature recorded during July does not change with depth across all lakes.
Ha: The mean lake temperature recorded during July change with depth across all lakes.


4.  Wrangle your NTL-LTER dataset with a pipe function so that the records meet the following criteria: 
 * Only dates in July. 
 * Only the columns: `lakename`, `year4`, `daynum`, `depth`, `temperature_C`
 * Only complete cases (i.e., remove NAs)

5. Visualize the relationship among the two continuous variables with a scatter plot of temperature by depth. Add a smoothed line showing the linear model, and limit temperature values from 0 to 35 °C. Make this plot look pretty and easy to read.

```{r scatterplot}
#4

cleanChemPhy <- 
  ChemPhy %>%
  mutate(month = month(sampledate)) %>%
  filter(month == "7") %>%
  select(lakename, 
         year4, 
         daynum, 
         depth, 
         temperature_C) %>%
  na.omit()

#5
TempDep <- ggplot(cleanChemPhy, 
       aes(x = depth, 
           y = temperature_C, 
           color = lakename))+
  geom_point(alpha = 1, 
             size = 1)+
  ylim(0,35)+
  labs(x = "Depth (m)", 
       y = "Temperature (C)", 
       color = "Lake Name") +
  geom_smooth(method = lm)
print(TempDep)
  
```


6. Interpret the figure. What does it suggest with regards to the response of temperature to depth? Do the distribution of points suggest about anything about the linearity of this trend?

> Answer: the figure suggests that temperature decreases as the depth increases. However, the distribution of the points suggests that the relationship between temperature and depths is not linear. the temperature drop was first slow, then fast, then slow again as the depth increases.


7. Perform a linear regression to test the relationship and display the results

```{r linear.regression}
#7
TempDepRegression <- lm(data = cleanChemPhy, temperature_C ~ depth)
summary(TempDepRegression)

```


8. Interpret your model results in words. Include how much of the variability in temperature is explained by changes in depth, the degrees of freedom on which this finding is based, and the statistical significance of the result. Also mention how much temperature is predicted to change for every 1m change in depth. 

> Answer: 73.87% of the variability in temperatre is explained by changes in depth, with the degree of freedom at 9726. The result is significant with p < 0.001 (p = 2.2 e-16). The result indicates that per 1m deeper into the lake, the temperature drops around 1.95 degree C.


---

## Multiple regression
Let's tackle a similar question from a different approach. Here, we want to explore what might the best set of predictors for lake temperature in July across the monitoring period at the North Temperate Lakes LTER. 


9. Run an AIC to determine what set of explanatory variables (year4, daynum, depth) is best suited to predict temperature.

10. Run a multiple regression on the recommended set of variables. 

```{r temperature.model}
#9
Temp.AIC <- lm (data = cleanChemPhy, 
             temperature_C ~ year4 + 
               daynum +
               depth)
step(Temp.AIC)


#10
Temp.model <- lm (data = cleanChemPhy, 
             temperature_C ~ year4 + 
               daynum +
               depth)
summary(Temp.model)
```

11. What is the final set of explanatory variables that the AIC method suggests we use to predict temperature in our multiple regression? How much of the observed variance does this model explain? Is this an improvement over the model using only depth as the explanatory variable?

> Answer: AIC suggests that all three explanatory variables (year4, daynum, depth) are needed to predict temperature. This modle could explain 74.12% of temperature change and is slightly better than only using depth.



---
## Analysis of Variance

12. Now we want to see whether the different lakes have, on average, different temperatures in the month of July. Run an ANOVA test to complete this analysis. (No need to test assumptions of normality or similar variances.) Create two sets of models: one expressed as an ANOVA models and another expressed as a linear model (as done in our lessons).

```{r anova.model}
#12
# As ANOVA
TempLakename.anova <- aov(data = cleanChemPhy, temperature_C ~ lakename)
summary(TempLakename.anova)

# As linear
TempLakename.anova2 <- lm(data = cleanChemPhy, temperature_C ~ lakename)
summary(TempLakename.anova2)
```

13. Is there a significant difference in mean temperature among the lakes? Report your findings. 

> Answer: there is significant difference in mean temperature among the lake because the p-value produced by ANOVA is less than 0.001



14. Create a graph that depicts temperature by depth, with a separate color for each lake. Add a geom_smooth (method = "lm", se = FALSE) for each lake. Make your points 50 % transparent. Adjust your y axis limits to go from 0 to 35 degrees. Clean up your graph to make it pretty. 

```{r scatterplot.2}
#14.

TempDep2 <- ggplot(cleanChemPhy, 
       aes(x = depth, 
           y = temperature_C, 
           color = lakename))+
  geom_point(alpha = 0.5, 
             size = 1)+
  ylim(0,35)+
  labs(x = "Depth (m)", 
       y = "Temperature (C)", 
       color = "Lake Name") +
  geom_smooth(method = "lm", se = FALSE)
print(TempDep)

```

15. Use the Tukey's HSD test to determine which lakes have different means.

```{r tukey.test}
#15

TukeyHSD(TempLakename.anova)
lakepair <- HSD.test(TempLakename.anova, "lakename", group = TRUE)
lakepair


```

16.From the findings above, which lakes have the same mean temperature, statistically speaking, as Peter Lake? Does any lake have a mean temperature that is statistically distinct from all the other lakes?

>Answer: statistically, Paul Lake and Ward Lake have the same mean temperature as Peter Lake. None of the lakes have a mean temperature that is statistically distinct from all the other lakes

 

17. If we were just looking at Peter Lake and Paul Lake. What's another test we might explore to see whether they have distinct mean temperatures? 

>Answer: t-test, chi-square


